---
title: "Overview and Trends"
author: "Data Science Group 31 - Ngoc Duong, Sabrina Lin, James Ng, Adeline Shin"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```
<p>

</p>

## Project overview
<p>

  We limited our data to taxis going into manhattan due to the size of the data available. Our logic is that this also reflects 
  
  that the borogh with the most people and most things to do and events on valentines day would be in manhattan. After filtering our data we still had to take a fraction of the data for each taxi line (10% for yellow and for hire cabs, 20% for green taxis) as the data was still too large to manage.
</p>


## General Trends

```{r ,echo = FALSE, warning = FALSE, message=FALSE}
vday_green_taxi <- read_csv("./data/vday_green_taxi.csv")
vday_yellow_taxi <- read_csv("./data/vday_yellow_taxi.csv")
vday_for_hire_vehicle <- read_csv("./data/vday_for_hire_vehicle.csv")

vday_for_hire_vehicle <- vday_for_hire_vehicle %>% 
  filter(hvfhs_license_num %in% c("HV0003", "HV0005"))

yellow_taxi_vday_samp <- sample_frac(vday_yellow_taxi, size = 0.1) %>% 
  mutate(type = "yellow")
green_taxi_vday_samp <- sample_frac(vday_green_taxi, size = 0.2) %>% 
  mutate(type = "green")
for_hire_vday_samp <- sample_frac(vday_for_hire_vehicle, size = 0.1) %>% 
  mutate(type = "for hire")

zone = read_csv("./data/taxi_zones.csv")

transport = bind_rows(yellow_taxi_vday_samp, green_taxi_vday_samp, for_hire_vday_samp) 

transport_final = left_join(transport, zone, by = c("pu_location_id" = "zone_id")) %>% 
  rename(pu_neighborhood = zone,
         pu_boro = borough) %>% 
  left_join(., zone, 
            by = c("do_location_id" = "zone_id")) %>%
  rename(do_neighborhood = zone,
         do_boro = borough) %>% 
  select(-ends_with("location_id")) %>% 
  filter(do_boro == "Manhattan") %>% 
  mutate(
  pu_time = paste(pickup_date, pickup_time, sep = " "),
  do_time = paste(dropoff_date, dropoff_time, sep = " "),
  duration = as.numeric(difftime(do_time, pu_time, units = "mins"))
) %>% 
  select(-do_time, -pu_time, -do_boro)

congestion = transport_final %>% 
  filter(total_amount>0) %>% 
  mutate(
    congestion = (fare_amount - ((trip_distance *.5/.2) +2.5))/.5
  ) %>% 
  filter(congestion>=0) %>% 
  mutate(
    percent_con = congestion/duration
  ) %>% 
  separate(
    pickup_time,
    c("pu_hour","pu_min","pu_sec"),
    sep = ":"
  ) %>% 
  mutate(
    pu_hour = as.numeric(pu_hour)
  ) %>% 
  filter(ratecode_id==1) %>% 
  filter(percent_con<1)
```

```{r ,echo = FALSE, warning = FALSE, message=FALSE}
congestion %>%  
  group_by(pu_hour, pu_boro) %>%
  filter(pu_boro != "NA", pu_boro != "Staten Island") %>% 
  summarize(avg_cong = mean(percent_con)) %>% 
  ggplot(aes(x=pu_hour, y = avg_cong))+
    geom_line()+
  facet_grid(~pu_boro)+
  geom_smooth(method = lm)+
  labs(
    title = "Average Duration of your Trip Stuck in Traffic When Traveling to Manhattan")+
    xlab("Pick Up Hour")+
    ylab("Average Congestion of Travel")
```

<p>

The trends show that congestion in manhattan is pretty consistent whereas it fluctuates more in the other boroughs when traveling into manhattan. 

</p>