---
title: "clean_analysis"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(tidyverse)
library(ggplot2)
set.seed(1)
```

Importing data

Reimporting data from csv (trying to make sure that it works)

```{r}
vday_green_taxi <- read_csv("./data/vday_green_taxi.csv")
vday_yellow_taxi <- read_csv("./data/vday_yellow_taxi.csv")
vday_for_hire_vehicle <- read_csv("./data/vday_for_hire_vehicle.csv")
```

# Sampling datasets

```{r}
yellow_taxi_vday_samp <- sample_frac(vday_yellow_taxi, size = 0.1) %>% 
  mutate(type = "yellow")
green_taxi_vday_samp <- sample_frac(vday_green_taxi, size = 0.2) %>% 
  mutate(type = "green")
for_hire_vday_samp <- sample_frac(vday_for_hire_vehicle, size = 0.1) %>% 
  mutate(type = "for hire")
```

Merge datasets
```{r}
zone = read_csv("./data/taxi_zones.csv")

transport = bind_rows(yellow_taxi_vday_samp, green_taxi_vday_samp, for_hire_vday_samp) 

transport_final = left_join(transport, zone, by = c("pu_location_id" = "zone_id")) %>% 
  rename(pu_neighborhood = zone,
         pu_boro = borough) %>% 
  left_join(., zone, 
            by = c("do_location_id" = "zone_id")) %>%
  rename(do_neighborhood = zone,
         do_boro = borough) %>% 
  select(-ends_with("location_id")) %>% 
  filter(do_boro == "Manhattan") 
```

Add duration variable
```{r}
transport_final = transport_final %>% 
  mutate(
  pu_time = paste(pickup_date, pickup_time, sep = " "),
  do_time = paste(dropoff_date, dropoff_time, sep = " "),
  duration = as.numeric(difftime(do_time, pu_time, units = "mins"))
) %>% 
  select(-do_time, -pu_time, -do_boro)
```

Probing for outliers (potentially false data records)
```{r eval = FALSE}
test = transport_final %>% 
  filter(duration > 100 & duration == 0) 
```

Eyeballing (outlier) test dataset, we decided to get rid of duration = 0 and duration > 150

```{r, echo = FALSE}
transport_final = transport_final %>% 
  filter(duration < 150 & duration != 0)
```

Probing into how many NA's are there (Bronx and Staten Island pick-up locations) 
```{r eval = FALSE}
length(transport_final$pu_neighborhood[is.na(transport_final$pu_neighborhood)])

# There are 1613 pickups happening in the Bronx and Staten Island
```


Make a heatmap for tips based on do and pu neighborhoods
```{r}
library("gplots")
#prepare data
tp_heatmap_tip = 
  transport_final %>% 
  filter(pu_boro == "Manhattan") %>% 
  select(do_neighborhood, pu_neighborhood, tip_amount) %>% 
  group_by(do_neighborhood, pu_neighborhood) %>% 
  summarize(avg_tip = mean(tip_amount)) %>% 
  ungroup() %>% 
  distinct() %>% 
  pivot_wider(
    names_from = pu_neighborhood,
    values_from = avg_tip
  ) %>% 
  mutate_if(is.numeric , replace_na, replace = 0)

m_tip <- as.matrix(tp_heatmap_tip[, -1])
rownames(m_tip) <- tp_heatmap_tip$do_neighborhood
heatmap.2(m_tip, scale="none", density.info="none", trace="none",  hclustfun = hclust)
```

Make a heatmap for count of drop-offs for drop-off and pick-up locations in manhattan
```{r}
tp_heatmap_count = 
  transport_final %>% 
  filter(pu_boro == "Manhattan") %>% 
  select(do_neighborhood, pu_neighborhood) %>% 
  group_by(do_neighborhood, pu_neighborhood) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  filter(count > 20) %>% 
  pivot_wider(
    names_from = pu_neighborhood,
    values_from = count
  ) %>% 
  mutate_if(is.numeric , replace_na, replace = 0)

m <- as.matrix(tp_heatmap_count[, -1])
rownames(m) <- tp_heatmap_count$do_neighborhood
heatmap.2(m, scale="none", density.info="none", trace="none",  hclustfun = hclust)
```

Load necessary packages for network building
```{r}
library(GGally)
library(network)
library(geomnet)
```

Try to build a network with transport data 
```{r, warning= FALSE, message= FALSE}
transport_final_mini =
  transport_final %>% 
  sample_frac(0.2) 

tp_list = 
  list("edges" = transport_final_mini[c("do_neighborhood", "pu_neighborhood")] %>% drop_na(),
       "vertices" = transport_final_mini[c("do_neighborhood", "pu_boro")] %>% drop_na()) 

tp.net <- network(tp_list$edges[, 1:2], directed = FALSE)

# create node attribute
tp.net %v% "do_neighborhood" <- as.character(
tp_list$vertices[ network.vertex.names(tp.net), "do_neighborhood"]
)
# create plot for ggnet2
set.seed(100)
ggnet2(tp.net, labelon = TRUE, 
size = 1, vjust = -0.6, mode = "kamadakawai", label.size = 1)
```

