---
title: "clean_analysis"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(tidyverse)
library(ggplot2)
set.seed(1)
```

# Importing data

```{r}
vday_green_taxi <- read_csv("./data/vday_green_taxi.csv")
vday_yellow_taxi <- read_csv("./data/vday_yellow_taxi.csv")
vday_for_hire_vehicle <- read_csv("./data/vday_for_hire_vehicle.csv")
```

# Filtering for Uber and Lyft Post Cleaned Data 

```{r}
vday_for_hire_vehicle <- vday_for_hire_vehicle %>% 
  filter(hvfhs_license_num %in% c("HV0003", "HV0005"))
```

# Sampling datasets

```{r}
yellow_taxi_vday_samp <- sample_frac(vday_yellow_taxi, size = 0.1) %>% 
  mutate(type = "yellow")
green_taxi_vday_samp <- sample_frac(vday_green_taxi, size = 0.2) %>% 
  mutate(type = "green")
for_hire_vday_samp <- sample_frac(vday_for_hire_vehicle, size = 0.1) %>% 
  mutate(type = "for hire")
```

# Merge datasets
```{r}
zone = read_csv("./data/taxi_zones.csv")

transport = bind_rows(yellow_taxi_vday_samp, green_taxi_vday_samp, for_hire_vday_samp) 

transport_final = left_join(transport, zone, by = c("pu_location_id" = "zone_id")) %>% 
  rename(pu_neighborhood = zone,
         pu_boro = borough) %>% 
  left_join(., zone, 
            by = c("do_location_id" = "zone_id")) %>%
  rename(do_neighborhood = zone,
         do_boro = borough) %>% 
  select(-ends_with("location_id")) %>% 
  filter(do_boro == "Manhattan") 
```

# Add duration variable
```{r}
transport_final = transport_final %>% 
  mutate(
  pu_time = paste(pickup_date, pickup_time, sep = " "),
  do_time = paste(dropoff_date, dropoff_time, sep = " "),
  duration = as.numeric(difftime(do_time, pu_time, units = "mins"))
) %>% 
  select(-do_time, -pu_time, -do_boro)
```

# Probing for outliers (potentially false data records)
```{r eval = FALSE}
test = transport_final %>% 
  filter(duration > 100 & duration == 0) 
```

# Eyeballing (outlier) test dataset, we decided to get rid of duration = 0 and duration > 150

```{r, echo = FALSE}
transport_final = transport_final %>% 
  filter(duration < 150 & duration != 0)
```

# Probing into how many NA's are there (Bronx and Staten Island pick-up locations) 
```{r eval = FALSE}
length(transport_final$pu_neighborhood[is.na(transport_final$pu_neighborhood)])

# There are 1613 pickups happening in the Bronx and Staten Island
```

#bar graph for all pick-ups and drop-offs, colored by borough
```{r}
transport_final %>% 
  filter(pu_boro!= "EWR") %>% 
  select(pu_neighborhood, do_neighborhood, pu_boro) %>% 
  group_by(pu_neighborhood, do_neighborhood, pu_boro) %>% 
  drop_na() %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = do_neighborhood, y = count, fill = pu_boro)) + geom_col() +
  coord_flip() 
```

# Dinner time subset of transport_final
```{r}
dinner_time_transport <- transport_final %>% 
  filter(stringr::str_detect(pickup_time, '^18|^19|^20'))
```

## Late night subset of transport_final
```{r}
late_night_transport <- transport_final %>% 
  filter(
    stringr::str_detect(pickup_time, '^21|^22|^23')|
    pickup_date == "2019-02-15"
    )
```

