---
title: "Heatmaps"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(d3heatmap)
knitr::opts_chunk$set(echo = TRUE)
set.seed(1)
```
To further investigate where people go on Valentine's Day via taxi or for hire car services, we made several heat maps to breakdown where people from each borough were coming into Manhattan. Due to the lack of drop-offs in Manhattan from Staten Island, we do not include Staten Island rides in any of the heat maps. 

If you are interested in looking at the general heat map trends, click [here](overview.html)!

## How to Read The Heat Maps 

The zones represented by the rows are the drop off locations in Manhattan, and the zones represented by the columns are the pick up locations in the respective borough. Both are arranged alphabetically for ease of finding specific pick-up to drop-off locations. If the zone in the respective borough is not present in the heat map, there were no rides coming from or into that zone. 

To see how many trips occurred between the desired locations, hover over the grid with the mouse. To zoom in on a specific area of any heat map, drag the mouse over the desired area. To highlight a specific pick-up or drop-off location, click on the desired location name.  

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
transport_final = read.csv("./data/transport_final.csv")

dinner_time_transport <- transport_final %>% 
  filter(stringr::str_detect(pickup_time, '^18|^19|^20'))

late_night_transport <- transport_final %>% 
  filter(
    stringr::str_detect(pickup_time, '^21|^22|^23')|
    pickup_date == "2019-02-15"
    )
```


```{r heatmap codes,echo=FALSE,warning=FALSE,message=FALSE}
manhattan_matrix <- function(input_boro, input_df) {
  man_man_df <- input_df %>% 
  filter(
    pu_boro == input_boro
  ) %>% 
  select(
    pu_neiborhood, do_neiborhood
  ) %>% 
  group_by(pu_neiborhood, do_neiborhood) %>% 
  summarise(count = n()) %>% 
  pivot_wider(
    names_from = pu_neiborhood,
    values_from = count
  ) %>% 
  replace(is.na(.),0) %>% 
  arrange(do_neiborhood) 

man_man_matrix <- as.matrix(man_man_df[-1])
row.names(man_man_matrix) <- pull(man_man_df, do_neiborhood)  
return(man_man_matrix)
}

heatmap_function <- function(input_matrix) {
  into_man_heatmap <- d3heatmap(input_matrix, scale = "none", colors = "RdPu",  dendrogram = "none", Rowv = FALSE, Colv = FALSE) 
  return(into_man_heatmap)
}
```

# All of Valentine's Day

The first set of heat maps looks at all the rides on Valentine's Day, including midnight to 2 AM on February 15th to account for post dinner plans. 

## Manhattan to Manhattan

The number of rides from any single pick-up location in Manhattan to any single drop-off location in Manhattan ranged from 0 to 218 rides. 

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
man_man_matrix <- manhattan_matrix("Manhattan", transport_final)

heatmap_function(man_man_matrix)
```

The Upper East Side loves itself, as many of the rides originate and end in either the Upper East Side North or South. Central Harlem and Central Harlem North, along with East Harlem North and East Harlem South also practice this form of self-love. 

Pick-ups and drop-offs in Midtown Center, East, North, and South, Murray Hill, and Penn Station are popular among most of the Manhattan locations. 

## Brooklyn to Manhattan

The number of rides from any single pick-up location in Brooklyn to any single drop-off location in Manhattan ranged from 0 to 19 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
brooklyn_man_matrix <- manhattan_matrix("Brooklyn", transport_final)

heatmap_function(brooklyn_man_matrix)
```

People in Williamsburg North and South, Greenpoint, DUMBO/Vinegar Hill, Downtown Brooklyn/MetroTech, and Brooklyn Heights were more likely to make the trek to Manhattan via a motor vehicle service. Popular destinations from Brooklyn look to be East Village, the Lower East Side, and TriBeCa/Civic Center. 

## Queens to Manhattan

The number of rides from any single pick-up location in Queens to any single drop-off location in Manhattan ranged from 0 to 80 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
queens_man_matrix <- manhattan_matrix("Queens", transport_final)

heatmap_function(queens_man_matrix)
```

Unsurprisingly, the most active pick-up locations in Queens on Valentine's Day are LaGuardia Airport and JFK Airport. Although this could be a coincidence, we would like to think that people are flying into New York to spend a special Valentine's Day with their special someone. 

For both airports, the most popular drop-off location is Times Square/Theater District, with Midtown Center, East, North, and South, Clinton East, and Murray Hill also as popular drop-off spots. 

## Bronx to Manhattan

The number of rides from any single pick-up location in the Bronx to any single drop-off location in Manhattan ranged from 0 to 19 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
bronx_man_matrix <- manhattan_matrix("Bronx", transport_final) 

heatmap_function(bronx_man_matrix)
```

People from the Bronx that ventured to Manhattan tended to stay within Northern Manhattan, as Central Harlem North, East Harlem North, Inwood, Hamilton Heights, and Washington Heights North and South received the most drop-off traffic. 

# Dinner time on Valentine's Day

Especially since Valentine's Day in 2019 was on a Thursday, most Valentine's Day plans would be scheduled after work. Since many restaurants and businesses have Valentine's Day events or specials, we decided to look at Manhattan drop offs from 6 PM to 9 PM. 

The following heat maps only encompass the dinner time period, allocated between 6 PM to 9 PM. 

## Manhattan to Manhattan

The number of rides from any single pick-up location in Manhattan to any single drop-off location in Manhattan from 6 PM to 9 PM ranged from 0 to 41 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
man_man_dinner <- manhattan_matrix("Manhattan", dinner_time_transport)

heatmap_function(man_man_dinner)
```

During dinner time hours, there are a noticable number of rides starting and ending in Lincoln Square East and East Village. East Village as a drop-off location receives more attention from Manhattan pick-up locations during dinner time compared to the overall Valentine's Day heatmap. (Perhaps there are compelling prix fixe Valentine Day's dinner menus in the area.) 

Lincoln Square East as both a drop-off and pick-up location also is more popular under the dinner time hour constraint. Appreciating the arts especially on Valentine's Day is always a good option. 

## Brooklyn to Manhattan

The number of rides from any single pick-up location in Manhattan to any single drop-off location in Manhattan from 6 PM to 9 PM ranged from 0 to 5 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
brooklyn_man_dinner <- manhattan_matrix("Brooklyn", dinner_time_transport)

heatmap_function(brooklyn_man_dinner)
```

East Village is the most popular destination for rides hailing from Brooklyn. Maybe it is the prix fixe menu specials? Sadly, knowing for sure is outside the scope of our analysis. 

Those from Williamsburg North and South and DUMBO/Vinegar Hill are still more drawn toward Manhattan compared to other Brooklyn zones. 

## Queens to Manhattan

The number of rides from any single pick-up location in Manhattan to any single drop-off location in Manhattan from 6 PM to 9 PM ranged from 0 to 15 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
queens_man_dinner <- manhattan_matrix("Queens", dinner_time_transport)

heatmap_function(queens_man_dinner)
```

Aside from people traveling from LaGuardia and JFK Airports, people from Queens are not traveling to Manhattan for Valentine's Day dinner. Some zones that are present in the overall Valentine's Day heat map such as Bayside are no longer present in the heat map above. 

## Bronx to Manhattan

The number of rides from any single pick-up location in Manhattan to any single drop-off location in Manhattan from 6 PM to 9 PM ranged from 0 to 5 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
bronx_man_dinner <- manhattan_matrix("Bronx", dinner_time_transport) 

heatmap_function(bronx_man_dinner)
```

Similarily to Queens, certain zones received no love. No one from the Bronx traveled to Midtown East, and no one from Pelham Bay traveled to Manhanttan. Central Harlem, Central Harlem North, Inwood, and Washington Heights North and South still remain relatively popular destinations. 

# Post Dinner on Valentine's Day  

What if you were kept late for work and could not make dinner or were already out celebrating Valentine's Day with dinner looking to extend the night? Whether it's to grab a drink or more food, we figured people would continue to enjoy their night past dinner. 

The following heat maps encompass only the later night hours from 9 PM of Valentine's Day to 2 AM of February 15th. 

## Manhattan to Manhattan

The number of rides from any single pick-up location in Manhattan to any single drop-off location in Manhattan from 9 PM on Valentine's Day to 2 AM on February 15th ranged from 0 to 38 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
man_man_late <- manhattan_matrix("Manhattan", late_night_transport)

heatmap_function(man_man_late)
```

The later night is when the Upper East Side zones calm down (or tire out.) East Village pick-ups reign supreme, as do East village drop-offs. Clinton East also has more pick-ups relative to other Manhattan pick-up zones. Clinton East and West, and East Chelsea also experienced an up-tick in drop-offs. Might as well enjoy a drink, especially on Valentine's Day.

## Brooklyn to Manhattan

The number of rides from any single pick-up location in Manhattan to any single drop-off location in Manhattan from 9 PM on Valentine's Day to 2 AM on February 15th ranged from 0 to 9 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
brooklyn_man_late <- manhattan_matrix("Brooklyn", late_night_transport)

heatmap_function(brooklyn_man_late)
```

Only the people from Williamsburg North and South to Manhattan seem to brave going to Manhattan via vehicle service. Riders from East Williamsburg also do not want to miss out, so it seems. 

## Queens to Manhattan

The number of rides from any single pick-up location in Manhattan to any single drop-off location in Manhattan from 9 PM on Valentine's Day to 2 AM on February 15th ranged from 0 to 16 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
queens_man_late <- manhattan_matrix("Queens", late_night_transport)

heatmap_function(queens_man_late)
```
Even though LaGuardia and JFK Airports are still the most popular pick-up destinations, only LaGuardia to Murray Hill has more than 10 rides. Hopefully this means that people who meant to spend Valentine's Day with their person made it on time! 

In addition, no rides went to Inwood from Queens, and no rides from Briarwood/Jamaica Hills went to Manhattan.

## Bronx to Manhattan

The number of rides from any single pick-up location in Manhattan to any single drop-off location in Manhattan from 9 PM on Valentine's Day to 2 AM on February 15th ranged from 0 to 5 rides.

```{r ,echo=FALSE, warning=FALSE, message=FALSE}
bronx_man_late <- manhattan_matrix("Bronx", late_night_transport) 

heatmap_function(bronx_man_late)
```

Interestingly, rides during the night came from more places in the Bronx and went to more places in Manhattan than during dinner hours. Popular drop-off locations match the same popular drop-off locations as the overall heat map: Central Harlem North, East Harlem North, Inwood, Hamilton Heights, and Washington Heights North and South. 
