---
title: "Heat Map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(d3heatmap)
```

```{r}
transport_final <- read_csv("./data/transport_final.csv")
```


# Functions for heatmaps 

## Matrix function


```{r}
manhattan_matrix <- function(input_boro, input_df) {
  man_man_df <- input_df %>% 
  filter(
    pu_boro == input_boro
  ) %>% 
  select(
    pu_neighborhood, do_neighborhood
  ) %>% 
  group_by(pu_neighborhood, do_neighborhood) %>% 
  summarise(count = n()) %>% 
  pivot_wider(
    names_from = pu_neighborhood,
    values_from = count
  ) %>% 
  replace(is.na(.),0) %>% 
  arrange(do_neighborhood) 

man_man_matrix <- as.matrix(man_man_df[-1])
row.names(man_man_matrix) <- pull(man_man_df, do_neighborhood)  
return(man_man_matrix)
}
```

## Heatmap function

```{r}
heatmap_function <- function(input_matrix) {
  into_man_heatmap <- d3heatmap(input_matrix, scale = "none", colors = "RdPu",  dendrogram = "none", Rowv = FALSE, Colv = FALSE) 
  return(into_man_heatmap)
}
```

# All of Valentine's day

## Manhattan to Manhattan

```{r}
man_man_matrix <- manhattan_matrix("Manhattan", transport_final)

heatmap_function(man_man_matrix)
```


## Brooklyn to Manhattan

```{r}
brooklyn_man_matrix <- manhattan_matrix("Brooklyn", transport_final)

heatmap_function(brooklyn_man_matrix)
```

## Queens to Manhattan

```{r}
queens_man_matrix <- manhattan_matrix("Queens", transport_final)

heatmap_function(queens_man_matrix)
```

## Bronx to Manhattan

```{r}
bronx_man_matrix <- manhattan_matrix("Bronx", transport_final) 

heatmap_function(bronx_man_matrix)
```

## Staten Island to Manhattan

```{r}
staten_island_matrix <- manhattan_matrix("Staten Island", transport_final)

heatmap_function(staten_island_matrix)
```

# Dinner time Valentine's Day

## Manhattan to Manhattan

```{r}
man_man_dinner <- manhattan_matrix("Manhattan", dinner_time_transport)

heatmap_function(man_man_dinner)
```


## Brooklyn to Manhattan

```{r}
brooklyn_man_dinner <- manhattan_matrix("Brooklyn", dinner_time_transport)

heatmap_function(brooklyn_man_dinner)
```

## Queens to Manhattan

```{r}
queens_man_dinner <- manhattan_matrix("Queens", dinner_time_transport)

heatmap_function(queens_man_dinner)
```

## Bronx to Manhattan

```{r}
bronx_man_dinner <- manhattan_matrix("Bronx", dinner_time_transport) 

heatmap_function(bronx_man_dinner)
```

## Staten Island to Manhattan

```{r}
staten_island_dinner <- manhattan_matrix("Staten Island", dinner_time_transport)

heatmap_function(staten_island_dinner)
```

# Late Night 

## Manhattan to Manhattan

```{r}
man_man_late <- manhattan_matrix("Manhattan", late_night_transport)

heatmap_function(man_man_late)
```


## Brooklyn to Manhattan

```{r}
brooklyn_man_late <- manhattan_matrix("Brooklyn", late_night_transport)

heatmap_function(brooklyn_man_late)
```

## Queens to Manhattan

```{r}
queens_man_late <- manhattan_matrix("Queens", late_night_transport)

heatmap_function(queens_man_late)
```

## Bronx to Manhattan

```{r}
bronx_man_late <- manhattan_matrix("Bronx", late_night_transport) 

heatmap_function(bronx_man_late)
```

## Staten Island to Manhattan

```{r}
staten_island_late <- manhattan_matrix("Staten Island", late_night_transport)

heatmap_function(staten_island_late)
```


# Wait time heat maps
```{r}
congestion_data = 
  taxi_data %>% 
  separate(pickup_time, into = c("pickup_hr", "pickup_min", "pickup_sec"), sep = ":") %>% 
  separate(dropoff_time, into = c("dropoff_hr", "dropoff_min", "dropoff_sec"), sep = ":") %>% 
  select(-pickup_sec, -dropoff_sec) %>% 
  mutate(
    pickup_hr = as.numeric(pickup_hr),
    pickup_min = as.numeric(pickup_min),
    dropoff_hr = as.numeric(dropoff_hr),
    dropoff_min = as.numeric(dropoff_min)
  ) %>% 
mutate(congestion = 
         (fare_amount - (trip_distance*2.5) -2.5)/.5) %>% 
  filter(congestion>=0)%>% 
mutate(wait_pct = congestion/duration) %>% 
  group_by(type, pickup_hr, pu_neiborhood, do_neiborhood) %>%  
  summarize(avg_congestion = mean(congestion),
            avg_wait_pct = mean(wait_pct))
```

```{r}
congestion_matrix <- function(input_boro) {
  congestion_data = 
  transport_final %>% 
  separate(pickup_time, into = c("pickup_hr", "pickup_min", "pickup_sec"), sep = ":") %>% 
  separate(dropoff_time, into = c("dropoff_hr", "dropoff_min", "dropoff_sec"), sep = ":") %>% 
  select(-pickup_sec, -dropoff_sec) %>% 
  mutate(
    pickup_hr = as.numeric(pickup_hr),
    pickup_min = as.numeric(pickup_min),
    dropoff_hr = as.numeric(dropoff_hr),
    dropoff_min = as.numeric(dropoff_min)
  ) %>% 
mutate(congestion = 
         (fare_amount - (trip_distance*2.5) -2.5)/.5) %>% 
  filter(congestion>=0)%>% 
mutate(wait_pct = congestion/duration) %>% 
  group_by(type, pickup_hr, pu_neighborhood, do_neighborhood) %>%  
  summarize(avg_congestion = mean(congestion),
            avg_wait_pct = mean(wait_pct)) %>% 
    select()
    arrange(do_neighborhood) %>% 
    pivot_wider(
      names_from = pu_neighborhood,
      values_from = 
    )
  
man_man_congestion_matrix <- as.matrix(congestion_data[-1])
row.names(man_man_congestion_matrix) <- pull(congestion_data, do_neighborhood)  
return(man_man_congestion_matrix)  
}
```

## Manhattan to Manhattan 
```{r}
man_man_congestion_matrix <- congestion_matrix("Manhattan")

heatmap_function(man_man_congestion_matrix)
```

